### 2014-07-16 ApTranscriptome_TR.Rmd
Cleaning up GSEA enrichment analysis and clustering for ms

# save results to file
write.table(file = paste(resultsdir, "Ap-responsive-transcripts", Sys.Date(), ".txt", sep = ""), Ap.response.type, row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
```

@@

# T-test comparing standard deviation of expression between colonies
(t.thermbreadth <- t.test(Ar.int.sd$exp_sd, A22.int.sd$exp_sd, alternative = "two.sided"))
```

@@

# t-test to compare standard deviation of 'Bimodal' transcripts between colonies
(t.themsens <- t.test(Ar.bim.sd, A22.bim.sd))
```

@@

```{r plot_T_on, echo=FALSE, eval=TRUE}
# merge data for plotting
Ap.T_on <- rbind(A22.high.T_on, Ar.high.T_on, A22.low.T_on, Ar.low.T_on)

# low plot
png("T_on_low.png")
T_on_low_plot <- ggplot(rbind(A22.low.T_on, Ar.low.T_on), aes(x = T_on, y = ..density.., colour = colony, group = colony, fill = colony)) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) +
    labs(x = "Temperature of gene activation", y = "Density") +
    theme(axis.title = element_text(size = rel(2))) +
    theme(axis.text = element_text(size = rel(1.2))) +
    theme(legend.position = "none")
print(T_on_low_plot)
dev.off()

# high plot
png("T_on_high.png")
T_on_high_plot <- ggplot(rbind(A22.high.T_on, Ar.high.T_on), aes(x = T_on, y = ..density.., colour = colony, group = colony, fill = colony)) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) +
    labs(x = "Temperature of gene activation", y = "Density") +
    theme(axis.title = element_text(size = rel(2))) +
    theme(axis.text = element_text(size = rel(1.2))) +
    theme(legend.position = "none")
print(T_on_high_plot)
dev.off()

# combined plot
T_on_plot <- ggplot(Ap.T_on, aes(x = T_on, y = ..density.., colour = colony, group = colony, fill = colony)) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) + 
    labs(x = "Temperature of gene activation", y = "Density") +
    theme(axis.title = element_text(size = rel(2))) +
    theme(axis.text = element_text(size = rel(1.2))) +
    theme(legend.position = "none")
print(T_on_plot)
```

@@

# report items in each cluster
GSEAReportClusters(Ap.high.hclust, h = 1)
write.table(unlist(GSEAReportClusters(Ap.high.hclust, h = 0.6)), file = "Ap_high_hclust.txt")
```

@@

# report items in each cluster at height 2
(Ap.low.hclust.report <- GSEAReportClusters(Ap.low.hclust, h = 2))
```

@@

# report items in each cluster at height 1.5
(Ap.bim.hclust.report <- GSEAReportClusters(Ap.bim.hclust, h = 1.5))
```

@@

# report items in each cluster at height 1.5
(Ap.int.hclust.report <- GSEAReportClusters(Ap.int.hclust, h = 1.5))
```

@@

@@

@@

```{r gsea_A22_high, results = 'hide'}
# A22 'High' genes
A22.geneList.high <- rep(0, length = length(Ap.geneList))
names(A22.geneList.high) <- names(Ap.geneList)
A22.geneList.high[(which(names(A22.geneList.high) %in% A22.high.transcripts))] <- 1
A22.high.gsea <- gsea(genelist = A22.geneList.high, geneID2GO = geneID2GO)
```

Clustering of enriched GO terms for *A. picea* **High** genes.

@@

@@

# combine
Ap.setdiff.gsea <- rbind(A22.gsea, Ar.gsea)
# reorder
Ap.setdiff.gsea <- Ap.setdiff.gsea[,c("Species", "Type", "GO.ID", "Term", "Annotated", "Significant", "Expected", "parentchild")]
colnames(Ap.setdiff.gsea)[8] <- "P"

write.csv(Ap.setdiff.gsea, file = paste(resultsdir, "Ap_setdiff_GSEA_", Sys.Date(), ".csv", sep = ""), row.names = FALSE)
```

```{r gsea_setdiff_table, echo = FALSE, results = 'asis', eval = FALSE}
#pandoc.table(Ap.setdiff.gsea[ , 1:4], caption = "Enriched GO terms found in one species but not the other for each thermal-response category")
print(xtable(Ap.setdiff.gsea[ , 1:4], comment = getOption("xtable.comment", FALSE)))
```

```{r GOSemSim}
# similarity among enriched GO terms
A22.low.GO.term.sim <- termSim(A22.low.gsea$GO.ID, A22.low.gsea$GO.ID, ont = "BP", organism = "fly")
# distance matrix
dist.A22.low.GO.term.sim <- dist(A22.low.GO.term.sim)
plot(hclust(dist(A22.low.GO.term.sim)))

A22.high.GO.term.sim <- termSim(A22.high.gsea$GO.ID, A22.high.gsea$GO.ID, ont = "BP", organism = "fly")
# distance matrix
plot(hclust(dist(A22.high.GO.term.sim)))
```

@@

```{r shiny_file, eval=FALSE, echo=FALSE}
shiny.dt <- TPM.dt.sub[Ap.response.type$Transcript]

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(RxNpval, sigtable, chi1, resp.TPM.dt.sub, Ap.type.table, tt2, tt3, qlslope, qlmean, GO.stress.Xsq, GO0006950.responsive, Ap.df, w1, w2, w3, Ap.T_on, Ap.BP.resultParentChild, Ap.gsea1, Ap.setdiff.gsea, resp.stress.responsive.count, t.mean, t.slope, t.curvature, t.wiggle, t.themsens, t.thermbreadth, file = "ms.RData")
```

>>>>

# save results to file
write.table(file = paste(resultsdir, "Ap_responsive_transcripts_", Sys.Date(), ".txt", sep = ""), Ap.response.type, row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
```

@@

# T-test comparing standard deviation of expression between colonies
(t.varint <- t.test(Ar.int.sd$exp_sd, A22.int.sd$exp_sd, alternative = "two.sided"))
```

@@

# t-test to compare standard deviation of 'Bimodal' transcripts between colonies
(t.varbim <- t.test(Ar.bim.sd, A22.bim.sd))
```

@@

```{r plot_T_on, echo=FALSE, eval=TRUE}
# merge data for plotting
Ap.T_on <- rbind(A22.high.T_on, Ar.high.T_on, A22.low.T_on, Ar.low.T_on)
Ap.T_on$type <- factor(Ap.T_on$type, levels = c("Low", "High"))
Ap.T_on$Species <- ifelse(Ap.T_on$colony == "ApVT", "A. picea", "A. carolinensis")

# combined plot
T_on_plot <- ggplot(Ap.T_on, aes(x = T_on, y = ..density.., colour = Species, group = Species, fill = Species)) +
    facet_grid(. ~ type) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) + 
    labs(x = "Temperature of gene activation", y = "Density") +
    theme(axis.title = element_text(size = rel(2))) +
    theme(axis.text = element_text(size = rel(1.2)))
#    theme(legend.position = "none")
print(T_on_plot)
```

@@

# report items in each cluster
Ap.high.hclust.report <- GSEAReportClusters(Ap.high.hclust, h = 1)

# add cluster information
Ap.high.gsea$Cluster <- NA

for(i in 1:length(Ap.high.hclust.report)) {
    Ap.high.gsea[Ap.high.hclust.report[[i]], "Cluster"] <- i
}
```

@@

# report items in each cluster
Ap.low.gsea.cluster <- Ap.low.hclust.report <- GSEAReportClusters(Ap.low.hclust, h = 1.5)

# add cluster information
Ap.low.gsea$Cluster <- NA

for(i in 1:length(Ap.low.hclust.report)) {
    Ap.low.gsea[Ap.low.hclust.report[[i]], "Cluster"] <- i
}
```

@@

# report items in each cluster
Ap.bim.hclust.report <- GSEAReportClusters(Ap.bim.hclust, h = 1.5)

# add cluster information
Ap.bim.gsea$Cluster <- NA

for(i in 1:length(Ap.bim.hclust.report)) {
    Ap.bim.gsea[Ap.bim.hclust.report[[i]], "Cluster"] <- i
}
```

@@

# report items in each cluster
(Ap.int.hclust.report <- GSEAReportClusters(Ap.int.hclust, h = 1.5))

# add cluster information
Ap.int.gsea$Cluster <- NA

for(i in 1:length(Ap.int.hclust.report)) {
    Ap.int.gsea[Ap.int.hclust.report[[i]], "Cluster"] <- i
}
```

```{r Ap_gsea_intersect_save, echo = FALSE}
Ap.high.gsea$Type <- "High"
Ap.low.gsea$Type <- "Low"
Ap.bim.gsea$Type <- "Bimodal"
Ap.int.gsea$Type <- "Intermediate"
Ap.gsea.intersect <- rbind(Ap.low.gsea, Ap.int.gsea, Ap.high.gsea, Ap.bim.gsea)

Ap.gsea.intersect <- Ap.gsea.intersect[,c("Type", "GO.ID", "Term", "Cluster", "Annotated", "Significant", "Expected", "parentchild")]
colnames(Ap.gsea.intersect)[which(colnames(Ap.gsea.intersect) == "parentchild")] <- "P"

write.csv(Ap.gsea.intersect, file = paste(resultsdir, "Ap_gsea_intersect_", Sys.Date(), ".csv", sep = ""), row.names = FALSE)
```

GSEA for genes that are *Intermediate* in A. carolinensis and *Bimodal* in A. picea

@@

Cluster by semantic similarity

@@

GSEA for genes that are *Intermediate* in A. carolinensis and *Low* in A. picea

@@

```{r gsea_A22_high, results = 'hide'}
# A22 'High' genes
A22.geneList.high <- rep(0, length = length(Ap.geneList))
names(A22.geneList.high) <- names(Ap.geneList)
A22.geneList.high[(which(names(A22.geneList.high) %in% A22.high.transcripts))] <- 1
A22.high.gsea <- gsea(genelist = A22.geneList.high, geneID2GO = geneID2GO)
```
Clustering of enriched GO terms for *A. picea* **High** genes.

@@

Combine gene set enrichment analysis for each species and category into single table.

@@

# combine
Ap.gsea.union <- rbind(A22.gsea, Ar.gsea)
# reorder
Ap.gsea.union <- Ap.gsea.union[,c("Species", "Type", "GO.ID", "Term", "Annotated", "Significant", "Expected", "parentchild")]
colnames(Ap.gsea.union)[8] <- "P"

write.csv(Ap.gsea.union, file = paste(resultsdir, "Ap_gsea_union", Sys.Date(), ".csv", sep = ""), row.names = FALSE)
```

@@

```{r shiny_file, eval=FALSE, echo=FALSE}
shiny.dt <- TPM.dt.sub[responsive.lms.ann.type[,list(Transcript, best.hit.to.nr.1, GO.Biological.Process.1)]]

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(RxNpval, sigtable, chi1, resp.TPM.dt.sub, Ap.type.table, tt2, tt3, qlslope, qlmean, GO.stress.Xsq, GO0006950.responsive, Ap.df, w1, w2, w3, Ap.T_on, Ap.BP.resultParentChild, resp.stress.responsive.count, t.mean, t.slope, t.curvature, t.wiggle, t.varint, t.varbim, T_on.high.ttest, T_on.low.ttest, Ap.gsea.intersect, Ap.gsea.union, file = "ms.RData")
```
